\documentclass[11pt]{article}
% \def\hidesolutions{}
\input{header}
\begin{document}
\exsheet{14}{19}{December} % parameters are the number of the session and the day

\begin{exercise}
    Given the following functions over an interval $[0,1)$, 
    \begin{enumerate}[label=(\alph*)]
        \item $f(x) = x$
        \item $g(x) = x^2$
        \item $h(x) = e^x$
        \item $s(x) = sin(\pi x)$
    \end{enumerate}
    sketch their extension to 
    \begin{itemize}
        \item a periodic function with period $1$,
        \item an even periodic function with period $2$,
        \item an odd periodic function with period $2$.
    \end{itemize}
    and state the formulas for these extended functions over the interval $[-1,1]$.
\end{exercise}

\begin{solution}     
\end{solution}



\begin{exercise}
    Consider the function 
    \begin{align*}
        f : [0,1] \to \bbR, \quad x \mapsto x^{3}.
    \end{align*}
    Extend this to an odd function with period $T = 2$. Sketch the graph of that function from $-2$ to $2$.
    Compute its Fourier coefficients in standard form. Compute the complex Fourier coefficients.
\end{exercise}
\begin{solution}     
\end{solution}







\begin{exercise}
    Suppose that
    \begin{align*}
        f(x) = \left\{\begin{array}{cc} x+1 & \text{ if $-1 < x < 0$ }, \\  1-x & \text{ if $ 0 < x < 1$ }, \\  0 & \text{ otherwise. } \end{array}\right.
        \qquad 
        g(x) = \left\{\begin{array}{cc} \frac 1 2 & \text{ if $-1 < x < 1$ }, \\ 0 & \text{ otherwise. } \end{array}\right.
        \qquad 
        h(x) = |x|
        .
    \end{align*}
    Compute the convolutions $u(x) = (f \star g)(x)$ and $v(x) = (g \star g)(x)$ and $w(x) = (g \star h)(x)$.
\end{exercise}
\begin{solution} 
\end{solution}



\begin{exercise}
    Suppose that $f(x) = x^2$ and that
    \[
        g(x) = \left\{\begin{array}{cc} \frac 1 2 & \text{ if $-1 < x < 1$ }, \\ 0 & \text{ otherwise. } \end{array}\right.
        \qquad 
        h(x) = \left\{\begin{array}{cc} e^{-x} & \text{ if $x > 0$ }, \\ 0 & \text{ otherwise. } \end{array}\right.
    \]
    Compute the convolutions $u(x) = (f \star g)(x)$ and $v(x) = (f \star h)(x)$ .
\end{exercise}

\begin{solution}     
\end{solution}


\begin{exercise}
    We have discussed solutions to the differential equation 
    \[
        - \Delta u(x) + k^2 u(x) = e^{-|x|}, \quad x \in \mathbb R.
    \]
    \begin{itemize}
        \item 
        Verify that, in the case $k=1$, we have a solution 
        \[
            u(x) = \frac 1 2 (1+|x|) e^{-|x|} 
        \]
        Verify that every function of the form 
        \[
            v(x) = \frac 1 2 (1+|x|) e^{-|x|} + c_1 e^{-x} + c_2 e^{x} 
        \]
        is a solution. For which values of $c_1$ and $c_2$ does the function decay towards zero as $x$ goes to $\pm \infty$?
        %
        \item 
        Verify that, in the case $k \neq 1$, we have a solution 
        \[
            u(x) = - \frac{ e^{-k|x|} }{ 2k (k^2-1) } + \frac{ e^{-|x|} }{ k^2-1 }
        \]
        Verify that every function of the form 
        \[
            v(x) = - \frac{ e^{-k|x|} }{ 2k (k^2-1) } + \frac{ e^{-|x|} }{ k^2-1 } + c_1 e^{-kx} + c_2 e^{kx} 
        \]        
        is a solution.
    \end{itemize}
\end{exercise}
\begin{solution}     
\end{solution}







\begin{exercise}
    We want to find a solution to the boundary value problem 
    \begin{gather*}
        - \Delta u(x) + k^2 u(x) = x, \quad 0 < x < L,
        \\
        u(0) = 0, \quad u(L) = 0.
    \end{gather*}
    \begin{itemize}
        \item Extend the right-hand side $f(x) = x$ to an odd function with period $2L$ and compute its Fourier coefficients.
        \item Using these coefficients, find the Fourier series of the solution $u$. Verify that the boundary condition $u(0) = u(L) = 0$ is satisfied.
    \end{itemize}
\end{exercise}
\begin{solution}     
\end{solution}




\begin{exercise}[Fun with Neumann boundary conditions]
    Consider the Poisson problem with \textit{Neumann boundary conditions} over the interval $[a,b] = [0,1]$: 
    \begin{gather*}
        - u''(x) + k^2 u(x) = x - \frac 1 2, \quad a < x < b,
        \\
        u'(a) = 0, \quad u'(b) = 0,
    \end{gather*}
    for some $k \geq 0$.
    \begin{enumerate}[label=(\alph*)]
        \item
        Extend $f(x) = x - \frac 1 2$ to an \textit{even} function over the real line with period $2$.
        \item 
        Compute the Fourier coefficients of that even extension of $f$.
        \item 
        Find the Fourier series of a function that satisfies the above differential equation.
    \end{enumerate}
\end{exercise}
\begin{solution}     
\end{solution}

\begin{exercise}[Fun with periodic boundary conditions]
    Consider the Poisson problem with \textit{periodic boundary conditions} over the interval $[a,b] = [0,1]$: 
    \begin{gather*}
        - u''(x) + k^2 u(x) = x - \frac 1 2, \quad a < x < b,
        \\
        u(a) = u(b),
    \end{gather*}
    for some $k \geq 0$.
    \begin{enumerate}[label=(\alph*)]
        \item
        Extend $f(x) = x - \frac 1 2$ to a $1$-periodic function over the real line.
        \item 
        Compute the Fourier coefficients of that extension of $f$.
        \item 
        Find the Fourier series of a function that satisfies the above differential equation.
    \end{enumerate}
\end{exercise}
\begin{solution}     
\end{solution}



% \begin{exercise}
% \end{exercise}
% \begin{solution}     
% \end{solution}
% 
% \begin{exercise}
% \end{exercise}
% \begin{solution}     
% \end{solution}

\end{document}

\begin{exercise}
    Find the Laplace transform $F(z)$ of 
    \begin{align*}
        f : \bbR_{0}^{+} \to \bbR, \quad t \mapsto t^{2}.
    \end{align*}
\end{exercise}
\begin{solution}     
    We compute 
    \begin{align*}
        F(z)&=\int_0^{\infty} t^2 e^{-z t} d t=\left[-\frac{t^2}{z} e^{-z t}\right]_0^{\infty}+\int_0^{\infty} \frac{2 t}{z} e^{-z t} d t \\ &=\left[\frac{2 t}{-z^2} e^{-z t}\right]_0^{\infty}+\int_0^{\infty} \frac{2}{z^2} e^{-z t} d t \\ &=\left[-\frac{2}{z^3} e^{-z t}\right]_0^{\infty} \\ &=0+\frac{2}{z^3} \\ &=\frac{2}{z^3}
        .
    \end{align*}
    Hence 
    \begin{align*}
        F(z)=\frac{2}{z^3} \quad \text { for } \mathrm{R}(z)>0.
    \end{align*}
\end{solution}

\begin{exercise}[Fun with Neumann boundary conditions 2]
    Suppose we have got a Poisson problem with Neumann boundary conditions over the interval $[a,b] = [0,1]$. 
    \begin{gather*}
        - u''(x) = f(x), \quad a < x < b,
        \\
        u'(a) = h_a, \quad u'(b) = h_b.
    \end{gather*}
    We need the compatibility condition 
    \begin{gather*}
        \int_a^b f(x) dx + ( h_b - h_a ) = 0.
    \end{gather*}
    \begin{enumerate}[label=(\alph*)]
        \item
        Find a constant function $c^f : [a,b] \to \bbR$ such that $f - c_f$ has integral zero.
        \item
        Use this to split up the original problem into two subproblems: 
        one with homogeneous boundary conditions, one with constant right-hand side, such that both problems satisfy the compatibility condition.
    \end{enumerate}
\end{exercise}
\begin{solution}     
\end{solution}



