\documentclass[11pt]{article}
\def\hidesolutions{}
\input{header}
\begin{document}
\exsheet{5}{10}{October} % parameters are the number of the session and the day





\begin{exercise}
    Consider a curve $u$ in one-dimensional space with
    \begin{align*}
        u : [1,2] \rightarrow \bbR, \qquad t \mapsto t^2 + t
    \end{align*}
    Verify that the curve is simple, differentiable, and regular. Compute the curve integral $\int_u f \;dl$, where 
    \begin{align*}
        f : \bbR \rightarrow \bbR, \qquad x \rightarrow 3x^3
    \end{align*}
    is the scalar field.
\end{exercise}
\begin{solution}[Remark]
    This is just a fancy way of describing integration by substitution from Analysis I.
\end{solution}
\begin{solution}
    We first verify that the curve is simple, differentiable, and regular.
    \begin{itemize}
        \item 
        The curve is simple if it does not intersect itself. This means that $u(t_1) \neq u(t_2)$ for $t_1 \neq t_2$. 
        In this case, the curve $u(t) = t^2 + t$ is a strictly increasing function on the interval $[1,2]$.
        We conclude that the curve is simple. 
        \item 
        We see that the curve is differentiable because its only component is differentiable. 
        \item 
        The curve is regular because the derivative $\dot u(t) = 2t + 1$ is not zero over the interval $(1,2)$. 
    \end{itemize}
    The curve integral of $f$ over $\Gamma$is given by
    \[
        \int_u f \, dl = \int_1^2 f(u(t)) \, |\dot u(t)| \, dt 
        = 
        \int_1^2 3(u(t))^3 \, |2t+1| \, dt 
        = 
        \int_1^2 3(t^2 + t)^3 \, |2t+1| \, dt 
        .
    \]
    We have $|2t+1| = 2t+1$ over the interval $[1,2]$. Hence
    \[
        \int_u f \, dl = 3 \int_1^2 (t^2 + t)^3 (2t + 1) \, dt.
    \]
    This integral can be computed using standard methods of integration. 
    We could just expand the polynomial that we integrate. Instead, we use substitution:
    \begin{align}
        \int_u f \, dl 
        = 
        \frac 3 4 \int_1^2 4(t^2 + t)^3 (2t + 1) \, dt
        = 
        \frac 3 4 \int_1^2 \partial_t (t^2 + t)^4 \, dt
        = 
        \frac 3 4 (t^2 + t)^4 |_{t=1}^{t=2}
        = 
        \frac 3 4 \left( 6^4 - 2^4 \right)
        .
    \end{align}
    This simplifies to 
    \begin{align}
        \int_u f \, dl 
        = 
        \frac 3 4 \cdot 2^4 \left( 3^4 - 1 \right) = \frac 3 4 \cdot 16 \cdot 80 = 3 \cdot 320 = 960
        .
    \end{align}
\end{solution}


\begin{exercise}[vector analysis in 1D]
    Let $\Omega \subseteq (a,b)$ be an open interval in one-dimensional space.
    \begin{itemize}
     \item Explain why there cannot be a simple closed continuous curve in $\Omega$.
     \item When $\Omega = (-10,10)$, compute the integral of the scalar field 
     \begin{gather*}
        f(x) = \frac{x}{\sqrt{1+x^2}}
     \end{gather*}
     along the curves 
     \begin{gather*}
        \gamma_1 : [0,1] \rightarrow \Omega, \quad t \mapsto (2t - 1),
        \\
        \gamma_2 : [-1,1] \rightarrow \Omega, \quad t \mapsto (t),
        \\
        \gamma_3 : [0,1] \rightarrow \Omega, \quad t \mapsto (1-2t),
        \\
        \gamma_4 : [0,1] \rightarrow \Omega, \quad t \mapsto (-1 + 2t^5),
    \end{gather*}
    Compute the tangent vectors $\dot\gamma(t)$.
    \item 
    Compute the integral of the vector field 
    \begin{align}
        F(x) = \left( x e^{x^2} \right)
    \end{align}
    along the curve $\gamma_4$. Find a potential for this vector field, and write down the general form of all potentials.
    \end{itemize}
\end{exercise}
\begin{solution}
    This is content of Analysis 1 but repackaged in the manner of vector analysis. 
    \begin{itemize}
        \item Any closed continuous curve in $\Omega$ would have to intersect itself in virtue of the Intermediate Value Theorem.
        \item Before we compute all the line integrals, we note that the curves $\gamma_1, \gamma_3$, and $\gamma_4$ map the interval $[0, 1]$ to the interval $[-1, 1]$.
        Hence, we can solve the corresponding line integrals by a change of variables in the line integral for $\gamma_2$.
        Furthermore, we note that $\gamma_3$ is the reparameterization of $\gamma_1$ in the opposite direction. Hence, we conclude from Exercise 5 on Exercise Sheet 4,
        that the corresponding line integrals must be equal. Let us now start the computations with $\gamma_2$:
        \begin{align}
            \int_{\gamma_2} f \;dl 
            &= 
            \int_{-1}^1 f(t) \;dt 
            = 
            \int_{-1}^1 \frac{t}{\sqrt{1+t^2}} \;dt 
            = 0,
        \end{align}
        since the integrand is odd around $t = 0$. For $\gamma_1$, we have
        \begin{align}
            \int_{\gamma_1} f \;dl 
            &= 
            \int_{0}^1 f(\gamma_1(t)) |\dot \gamma_1(t)| \;dt 
            = 
            2\int_{0}^1 f(2t - 1) \;dt 
            = \int_{-1}^1 f(s) \;ds = 0 = \int_{\gamma_2} f \;dl = 0.
        \end{align}
        As already pointed out previously, we must have $\int_{\gamma_3} f \;dl = \int_{\gamma_1} f \;dl = 0$. 
        Finally, we find for $\gamma_4$ that $\dot \gamma_4(t) = 10t^4 \geq 0$ for $t \in [0,1]$
        and thus
        \begin{align}
            \int_{\gamma_4} f \;dl 
            = 
            \int_{0}^1 f(\gamma_4(t)) \dot \gamma_4(t) \;dt 
            = 
            \int_{-1}^1 f(u) \;du = \int_{\gamma_2} f \;dl = 0,
        \end{align}
        where we have used the substitution $u = \gamma_4(t) = -1 + 2t^5$ with $du = \dot \gamma_4(t) \;dt$.
        \item The general form of a potential for $F$ is $f(x) = \frac 1 2 e^{x^2} + C$, where $C$ is an arbitrary constant. We set
        $C = -1/2$ such that $f(0) = 0$. For the curve integral of $F$ along $\gamma_4$, we thus find that
        \begin{align}
            \int_{\gamma_4} F dl = f(\gamma_4(1)) - f(\gamma_4(0)) = f(1) - f(-1) = \frac 1 2 (e - e^{-1}).
        \end{align}
    \end{itemize}
\end{solution}

\begin{exercise}
    We review notions of potentials and conservative vector fields. Let $\Omega \subseteq \bbR^n$ be open.
    Suppose we have a vector field $F = (F_1,\dots,F_n) \in C^1(\Omega,\bbR^n)$.
	Recall that we have introduced the condition
    \begin{align}
    \label{eq:conservative}
        \partial_i F_j = \partial_j F_i, \qquad 1 \leq i,j \leq n.
    \end{align}
    \begin{itemize}
     \item Suppose that $n=2$. Show that $F$ satisfies \eqref{eq:conservative} if and only if it is curl-free: $\operatorname{curl} F = 0$.
     \item Suppose that $n=3$. Show that $F$ satisfies \eqref{eq:conservative} if and only if it is curl-free: $\operatorname{curl} F = 0$.
     \item Suppose that $n=1$. Show that $F$ satisfies \eqref{eq:conservative}.
     \item Suppose that $F$ admits a potential $f \in C^1(\Omega,\bbR)$, so that $\nabla f = F$. 
     Show that if $\gamma : [a,b] \rightarrow \Omega$ is a simple regular curve, then 
     \begin{align}
        \int_\gamma F \;dl = f(\gamma(b)) - f(\gamma(a)).
     \end{align}
     Show that if $\gamma$ is closed, then 
     \begin{align}
        \int_\gamma F \;dl = 0.
     \end{align}
    \end{itemize}
\end{exercise}
\begin{solution}
\begin{itemize}
    \item We recall that the curl of a vector field $F = (F_1,F_2)$ is given by
    \begin{align}
        \operatorname{curl} F = \left( \partial_1 F_2 - \partial_2 F_1 \right).
    \end{align}
    Hence, it is clear that $\operatorname{curl} F = 0$ if and only if $F$ satisfies \eqref{eq:conservative}.
    \item We recall that the curl of a vector field $F = (F_1,F_2,F_3)$ is given by
    \begin{align}
        \operatorname{curl} F = 
        \begin{pmatrix}
            \partial_2 F_3 - \partial_3 F_2
            \\
            \partial_3 F_1 - \partial_1 F_3
            \\
            \partial_1 F_2 - \partial_2 F_1
        \end{pmatrix}.
    \end{align}
    Hence, it is clear that $\operatorname{curl} F = 0$ if and only if $F$ satisfies \eqref{eq:conservative}. 
    Also, recall Exercise 6 from Exercise Sheet 3. On the one hand, we noticed there that the curl of a vector field is given by the off-diagonal entries of the
    anti-symmetric part of the Jacobian matrix. On the other hand, condition \eqref{eq:conservative} is equivalent to the statement that the Jacobian matrix of $F$ is symmetric.
    We therefore conclude that the curl vanishes if and only if the Jacobian matrix is symmetric.
    \item For $n=1$, we have $F = (F_1)$ and the condition \eqref{eq:conservative} is trivially satisfied.
    \item We have $\nabla f = F$. Hence, we can write the line integral as
    \begin{align}
        \int_\gamma F \;dl &= \int_a^b F(\gamma(t)) \cdot \dot \gamma(t) \;dt \\
        &= \int_a^b \nabla f(\gamma(t)) \cdot \dot \gamma(t) \;dt = \int_a^b \frac{d}{dt} f(\gamma(t)) \;dt = f(\gamma(b)) - f(\gamma(a)).
    \end{align}
    If $\gamma$ is closed, then $\gamma(a) = \gamma(b)$ and we find that the line integral vanishes.
\end{itemize}
\end{solution}







\begin{exercise}
    We introduce the following curves:
    \begin{gather*}
        \gamma : [0,1] \rightarrow \bbR^{3}, \qquad t \mapsto \left( 3, t^2, 4t \right),
        \\
        \delta : [1,\infty) \rightarrow \bbR^{2}, \qquad t \mapsto \left( 5, e^{-t} \right)
    \end{gather*}
    For each curve
    \begin{itemize}
        \item compute the tangent vector
        \item compute the speed of the curve
        \item find the unit tangent vector
        \item find the unit normal along the curve that is the 90 degree clockwise rotation of unit tangent
        \item argue why it is a regular curve
        \item and compute the length of the curve.
    \end{itemize}
\end{exercise}
\begin{solution}
    \begin{itemize}
        \item We compute the tangent vectors:
        \begin{align}
            \dot \gamma(t) = \begin{pmatrix} 0 \\ 2t \\ 4 \end{pmatrix}, \quad \dot \delta(t) = \begin{pmatrix} 0 \\ -e^{-t} \end{pmatrix}.
        \end{align}
        \item The speed of the curve is given by the norm of the tangent vector:
        \begin{align}
            |\dot \gamma(t)| = \sqrt{4t^2 + 16} = 2 \sqrt{t^2 + 4}, \quad |\dot \delta(t)| = e^{-t}.
        \end{align}
        \item The unit tangent vector is given by
        \begin{align}
            \hat t_\gamma(t) &= \frac{\dot \gamma(t)}{|\dot \gamma(t)|}  = \frac{1}{\sqrt{t^2 + 4}} \begin{pmatrix} 0 \\ t \\ 2 \end{pmatrix},\\
            \hat t_\delta(t) &= \frac{\dot \delta(t)}{|\dot \delta(t)|} = \begin{pmatrix} 0 \\ -1 \end{pmatrix}.
        \end{align}
        Note that the unit tangent vector of $\delta$ is constant!
        \item The unit normal vector is given by the 90 degree clockwise rotation of the unit tangent vector. The corresponding rotation matrix is
        \begin{align}
            R = \begin{pmatrix} 0 & -1 \\ 1 & 0 \end{pmatrix}.
        \end{align}
        Hence, we find
        \begin{align}
            \hat n_\delta(t) = R \hat t_\delta(t) = \begin{pmatrix} 1 \\ 0 \end{pmatrix}.
        \end{align}
        \item The curves are regular because their tangent vectors never vanish.
        \item The length of the curves is given by the integral of the speed: For $\gamma$, we find
        \begin{align}
            \int_0^1 |\dot \gamma(t)| dt &= 2 \int_0^1 \sqrt{t^2 + 4} dt = 4 \int_{0}^1 \sqrt{(t/2)^2 + 1} dt\\
            &= 8 \int_0^{1/2} \sqrt{u^2 + 1} du =  4(\sqrt{2} + \sinh^{-1}(1/2)),\\
        \end{align}
        where we have used the substitution $u = t/2$ with $du = dt/2$. The last integral can be found with \emph{trigonometric substitution}.
        
        For $\delta$, we find
        \begin{align}
            \int_1^\infty |\dot \delta(t)| dt = \int_1^\infty e^{-t} dt = e^{-1}.
        \end{align}
        Note that the curve is parametrized for $t \in [1,\infty)$, but still has a finite length!
    \end{itemize}
\end{solution}



\begin{exercise}
	We once again consider the vector field 
	\[
		F : \bbR^2 \rightarrow \bbR^2, \quad (x,y) \mapsto \begin{pmatrix} \frac{y}{x^2+y^2} \\ \frac{-x}{x^2+y^2} \end{pmatrix}.
	\]
	We want to find a potential over the domain 
	\[
        \Omega = \left\{ (x,y) \in \bbR^2 : x > 0 \right\}
	\]
\end{exercise}
	\begin{itemize}
        \item 
        It is good to always begin with picture: make a short sketch of this domain
        \item 
        Our idea is that we construct a curve from $(1,0)$ to $(x,y)$ as follows:
        \begin{align}
            \gamma_{x,y} : [0,2] \rightarrow \bbR^2, \quad t \mapsto \left\{\begin{array}{ll}
                                                                \left( 1, ty \right) & t \in [0,1]
                                                                \\
                                                                \left( 1 + (t-1)(x-1),  y \right)  & t \in (1,2]
                                                               \end{array}\right.
        \end{align}
        Verify that this curve is piecewise differentiable and regular. 
        Make a drawing and understand the path of this curve.
        \item 
        We fix some constant of integration $C$ at $(0,1)$. Show that 
        \[
            f(x,y) = \int_{\gamma_{x,y}} F \;dl.
        \]
        is a scalar potential for this vector field.
	\end{itemize}
\begin{solution}
\end{solution}


\begin{exercise}
    The closed curve 
    \begin{align*}
        \gamma(t) = \left( \sin(t)(1+0.5 \sin(2t)), \cos(t)(1+0.5 \sin(2t)) \right)
    \end{align*}
    encircles a domain $\Omega$ in counterclockwise direction. 
	Find the tangent $\dot\gamma(t)$, the unit tangent $\tau(t)$ and the outward pointing unit normal $\vec n(t)$.
    Only simplify as much as reasonable. 
\end{exercise}
\begin{solution}
    We calculate:
    \begin{align*}
        \dot\gamma(t) 
        = \left( 
            \cos(t)( 1+0.5 \sin(2t) ) + \sin(t)\cos(2t)
            ,
            -\sin(t)( 1+0.5 \sin(2t) ) + \cos(t)\cos(2t)
        \right)
    \end{align*}
    With that:
    \begin{align*}
        |\dot\gamma(t)|^{2}
        &= 
        \cos(t)^{2}( 1+0.5 \sin(2t) )^{2} + \sin(t)^{2}\cos(2t)^{2} + 2 \cos(t)( 1+0.5 \sin(2t) ) \sin(t)\cos(2t)
        \\&\quad \quad 
        +
        \sin(t)^{2}( 1+0.5 \sin(2t) )^{2} + \cos(t)^{2}\cos(2t)^{2} - 2\sin(t)( 1+0.5 \sin(2t) ) \cos(t)\cos(2t)
        \\&= 
        ( 1+0.5 \sin(2t) )^{2} + \cos(2t)^{2} + 2 \cos(t)( 1+0.5 \sin(2t) ) \sin(t)\cos(2t)
        \\&\quad \quad 
        - 2 \sin(t)( 1+0.5 \sin(2t) ) \cos(t)\cos(2t)
        \\&= 
        ( 1+0.5 \sin(2t) )^{2} + \cos(2t)^{2}
        .
    \end{align*}
    Hence 
    \begin{align*}
        |\dot\gamma(t)|
        &= 
        \sqrt{ ( 1+0.5 \sin(2t) )^{2} + \cos(2t)^{2} }
        .
    \end{align*}
    We want to compute the tangent vector:
    \begin{align*}
        \tau(t) 
        = 
        \left( 
            \frac{\cos(t)( 1+0.5 \sin(2t) ) + \sin(t)\cos(2t)}{\sqrt{ ( 1+0.5 \sin(2t) )^{2} + \cos(2t)^{2} }}
            , 
            \frac{-\sin(t)( 1+0.5 \sin(2t) ) + \cos(t)\cos(2t)}{\sqrt{ ( 1+0.5 \sin(2t) )^{2} + \cos(2t)^{2} }} 
        \right)
    \end{align*}
    Accordingly, the normal vector is:
    \begin{align*}
        \vec n(t) 
        = 
        \left( 
            \frac{-\sin(t)( 1+0.5 \sin(2t) ) + \cos(t)\cos(2t)}{\sqrt{ ( 1+0.5 \sin(2t) )^{2} + \cos(2t)^{2} }} 
            , 
            - \frac{\cos(t)( 1+0.5 \sin(2t) ) + \sin(t)\cos(2t)}{\sqrt{ ( 1+0.5 \sin(2t) )^{2} + \cos(2t)^{2} }}
        \right)
    \end{align*}
    

    \begin{tikzpicture}
        \begin{axis}[
            domain=0:360,
            xmin=-2,
            xmax=2,
            ymin=-2,
            ymax=2,
            xlabel={x},
            ylabel={y}
        ]
        \addplot[blue, variable=\t, samples=100] ({sin(\t)*( 1 + 0.5 * sin(2*\t) )}, {cos(\t)*( 1 + 0.5 * sin(2*\t) )});
        \end{axis}
    \end{tikzpicture}
\end{solution}














\begin{exercise}
    We work over the quadratic domain 
    \begin{gather*}
        \Omega := \left\{ (x_1,x_2) \in \bbR^{2} \suchthat* -1 < x_1 < 1, \; -1 < x_2 < 1 \right\}.
    \end{gather*}
    Compute the integral $\iint_\Omega \divergence \vec F \;dx_1dx_2$, where 
    \begin{gather*}
        \vec F(x_1,x_2) = \left( \sin(x_1) x_2, \left( x_1^2 + x_2 \right)^5 \right)
    \end{gather*}
\end{exercise}
\begin{solution}
    % Express this volume integral as a curve integral.
    We make use of the Divergence theorem to express the volume integral as a curve integral.
    \begin{align*}
        \int \int_{\Omega} \nabla \cdot \vec{F} \;d x_1 \:d x_2 
        =
        &
        \int_{\partial \Omega}\vec{F}\cdot\vec{n} \;d\ell,
        \\&=
        \int_{-1}^1 \vec{F}(x_1,-1)\cdot\begin{pmatrix}0\\-1\end{pmatrix} \;dx_1 + \int_{-1}^1 \vec{F}(x_1,1)\cdot\begin{pmatrix}0\\1\end{pmatrix} \;dx_1,
        \\
        & + \int_{-1}^1 \vec{F}(-1,x_2)\cdot\begin{pmatrix}-1\\0\end{pmatrix} \;dx_2 + \int_{-1}^1 \vec{F}(1,x_2)\cdot\begin{pmatrix}1\\0\end{pmatrix} \;dx_2,s
        \\&=
        \int_{-1}^1 (1-x_1^ 2)^ 5 \;dx_1 + \int_{-1}^1 (x_1^2 + 1)^ 5 \;dx_1 + \int_{-1}^1 \sin(1)x_2\;dx_2 +  \int_{-1}^1 -\sin(-1)x_2\;dx_2,\\
    \end{align*}
    Since $x_2$ is an odd function, so the last two integrals are 0. To evaluate the first two integrals,
    we can either compute the quintic powers manually, which is a lot of work, or we utilize the binomial theorem:
    \begin{gather*}
        (x+y)^n=\sum_{k=0}^n\left(\begin{array}{l}
        n \\
        k
        \end{array}\right) x^{n-k} y^k
    \end{gather*}
    With that, we can simplify the calculation as follows: 
    \begin{align*}
        \int \int_{\Omega} \nabla \cdot \vec{F} \;d x_1 \:d x_2 
        &=
        \int_{\partial \Omega}\vec{F}\cdot\vec{n} \;d\ell,
        \\&=
        \int_{-1}^1 (1-x_1^2)^5 \;dx_1 + \int_{-1}^1 (x_1^2 + 1)^ 5 \;dx_1
        \\&=
        \sum_{k = 0}^5 \binom{5}{k} (-1)^{k} \int_{-1}^1 (x)^{2k} \;dx_1 + \sum_{k = 0}^5 \binom{5}{k}\int_{-1}^1 (x)^{2k} \;dx_1
        \\&=
        \sum_{k = 0}^5 \binom{5}{k} (-1)^{k} \left[\frac{x^{2k+1}}{2k+1}\right]_{-1}^1 + \sum_{k = 0}^5 \binom{5}{k} \left[\frac{x^{2k+1}}{2k+1}\right]_{-1}^1
        \\&=
        \sum_{k = 0}^5 \binom{5}{k} \left( (-1)^{k} + 1 \right) 
        \left[\frac{x^{2k+1}}{2k+1}\right]_{-1}^1
    \end{align*}
    For $k = 1, 3, 5$, we get $\left( (-1)^{k} + 1 \right) = 0$. 
    So we only need the terms with $k = 0, 2, 4$.
    Thus
    \begin{align*}
        &
        \sum_{k = 0}^5 \binom{5}{k} \left( (-1)^{k} + 1 \right) 
        \left[\frac{x^{2k+1}}{2k+1}\right]_{-1}^1
        \\&=
        2
        \binom{5}{0}
        \left[\frac{x^{0+1}}{0+1}\right]_{-1}^1
        +
        2
        \binom{5}{2}
        \left[\frac{x^{4+1}}{4+1}\right]_{-1}^1
        +
        2
        \binom{5}{4}
        \left[\frac{x^{8+1}}{8+1}\right]_{-1}^1
        \\&=
        2
        \binom{5}{0}
        \left[\frac{x^{1}}{1}\right]_{-1}^1
        +
        2
        \binom{5}{2}
        \left[\frac{x^{5}}{5}\right]_{-1}^1
        +
        2
        \binom{5}{4}
        \left[\frac{x^{9}}{9}\right]_{-1}^1
        \\&=
        4
        \binom{5}{0}
        +
        \frac 4 5
        \binom{5}{2}
        +
        \frac 4 9
        \binom{5}{4}
        \\&=
        \frac{128}{9}
        .
    \end{align*}
\end{solution}

















\end{document}
