\documentclass[11pt]{article}
\def\hidesolutions{}
\input{header}
% \usepackage{enumitem}
\begin{document}
\exsheet{13}{12}{December} % parameters are the number of the session and the day











\begin{exercise}
    The Fourier transform of 
    \begin{align*}
        f(x) = e^{-5x^2}
    \end{align*}
    is the function 
    \begin{align*}
        \hat f(x) = \frac{1}{\sqrt{10}} e^{-\frac{\alpha^2}{20}}.
    \end{align*}
    Find the Fourier transforms of 
    \begin{align*}
     f', \quad f'', \quad f''', \quad f'''', \quad g(x) = f(2x), \quad h(x) = f(x-3).
    \end{align*}
\end{exercise}
\begin{solution}  
    To evaluate the Fourier transform of the derivative of a function we will use the following identity from the lecture:
    \[
        \mathfrak{F}\left(f^{(n)}\right)(\alpha)=(i \alpha)^n \mathfrak{F}(f)(\alpha)
    \]
    We then find 
    \begin{itemize}
    \item $\mathfrak{F}\left(f' \right)(\alpha) = i \alpha \mathfrak{F}(f)(\alpha) = \frac{i \alpha}{\sqrt{10}} e^{-\frac{\alpha^2}{20}}$
    \item $\mathfrak{F}\left(f'' \right)(\alpha) = -\alpha^2 \mathfrak{F}(f)(\alpha) = -\frac{ \alpha^2}{\sqrt{10}} e^{-\frac{\alpha^2}{20}}$
    \item $\mathfrak{F}\left(f' \right)(\alpha) = -i \alpha^3 \mathfrak{F}(f)(\alpha) = \frac{-i \alpha^3}{\sqrt{10}} e^{-\frac{\alpha^2}{20}}$
    \item $\mathfrak{F}\left(f' \right)(\alpha) = \alpha^4 \mathfrak{F}(f)(\alpha) = \frac{ \alpha^4}{\sqrt{10}} e^{-\frac{\alpha^2}{20}}$
    \end{itemize}
    For the Fourier transform of $g(x)$ we use the following identity from the lecture: 
    $g(x)=e^{-i b x} f(a x)$ has the Fourier transform
    $\hat{g}(\alpha)=\frac{1}{|a|} \hat{f}\left(\frac{\alpha+b}{a}\right)$,
    consequently:
    \begin{align*}
        \mathfrak{F}(g(x))(\alpha) & =\mathfrak{F}(f(2 x))(\alpha) \\ & =\frac{1}{2} \mathfrak{F}(f(x))\left(\frac{\alpha}{2}\right) \\ & =\frac{1}{2 \sqrt{10}} e^{-\frac{\left(\frac{\alpha}{2}\right)^2}{20}} \\ & =\frac{1}{2 \sqrt{10}} e^{-\frac{\alpha^2}{80}}
    \end{align*}
    For the Fourier transform of $h(x)$ we use a change of variables:
    \begin{align*}
        \mathfrak{F}(h(x))(\alpha)
        &= 
        \frac{1}{\sqrt{2\pi}}
        \int_{-\infty}^{+\infty} e^{-5(x-3)^2} e^{ - \imath x \alpha } \;dx
        \\&= 
        \frac{1}{\sqrt{2\pi}}
        \int_{-\infty}^{+\infty} e^{-5x^2} e^{ - \imath (x + 3) \alpha } \;dx
        \\&= 
        \frac{1}{\sqrt{2\pi}}
        \int_{-\infty}^{+\infty} e^{-5x^2} e^{ - \imath x \alpha } e^{ - \imath 3 \alpha }\;dx
        \\&= 
        \frac{ e^{ - \imath 3 \alpha }}{\sqrt{2\pi}}
        \int_{-\infty}^{+\infty} e^{-5x^2} e^{ - \imath x \alpha }\;dx
        \\&= 
        e^{ - \imath 3 \alpha } 
        \mathfrak{F}(f(x))(\alpha)
        \\&= 
        \frac{e^{ - \imath 3 \alpha }}{\sqrt{10}} e^{-\frac{\alpha^2}{20}}.
        \\&= 
        \frac{1}{\sqrt{10}} e^{-\frac{\alpha^2}{20} - \imath 3 \alpha }
        .
    \end{align*}
\end{solution}




















\begin{exercise}
    Find the Laplace transform $F(z)$ of 
    \begin{align*}
        f : \bbR_{0}^{+} \to \bbR, \quad t \mapsto t^{2}.
    \end{align*}
\end{exercise}
\begin{solution}     
    We compute 
    \begin{align*}
        F(z)&=\int_0^{\infty} t^2 e^{-z t} d t=\left[-\frac{t^2}{z} e^{-z t}\right]_0^{\infty}+\int_0^{\infty} \frac{2 t}{z} e^{-z t} d t \\ &=\left[\frac{2 t}{-z^2} e^{-z t}\right]_0^{\infty}+\int_0^{\infty} \frac{2}{z^2} e^{-z t} d t \\ &=\left[-\frac{2}{z^3} e^{-z t}\right]_0^{\infty} \\ &=0+\frac{2}{z^3} \\ &=\frac{2}{z^3}
        .
    \end{align*}
    Hence 
    \begin{align*}
        F(z)=\frac{2}{z^3} \quad \text { for } \mathrm{R}(z)>0.
    \end{align*}
\end{solution}


















\begin{exercise}
    Find $f : \mathbb R \to \mathbb R$ such that 
    \begin{align*}
        \hat f(\alpha) = \frac{3}{1+\alpha^{2}} + \frac{-1}{1+4\alpha^{2}} + \frac{ \sin( 4 \alpha + 3 ) }{ 4 \alpha + 3 }
    \end{align*}
\end{exercise}
\begin{solution}   Because of the linearity property we can treat the inverse of each term in seperately. Therefore 
 \begin{itemize}
    \item $\begin{aligned} \mathfrak{F}^{-1}\left(\frac{\sin (4 \alpha+3)}{4 a r+3}\right)(x) & =\frac{\frac{1}{4}}{\frac{1}{4}} \mathfrak{F}^{-1}\left(\frac{\sin \left(\frac{\alpha+3 / 4}{\frac{1}{4}}\right)}{\frac{\alpha+\frac{3}{4}}{\frac{1}{4}}}\right)(x) \\ & =\frac{1}{4} e^{-i \frac{3}{4} x} \sqrt{\frac{\pi}{2}} \mathfrak{F}^{-1}\left(\sqrt{\frac{2}{\pi}} \frac{\sin (\alpha)}{\alpha}\right)\left(\frac{x}{4}\right) \\ & = \begin{cases}\frac{1}{4} \sqrt{\frac{\pi}{2}} e^{-i \frac{3}{4} x} & \text { if }-4 \leq x \leq 4 \\ 0 & \text { else }\end{cases} \end{aligned}$
    \item $\begin{aligned} \mathfrak{F}^{-1}\left(\frac{3}{1+\alpha^2}\right)(x) & =3 \sqrt{\frac{\pi}{2}} \mathfrak{F}^{-1}\left(\sqrt{\frac{2}{\pi}} \frac{1}{1+a^2}\right)(x) \\ & =3 \sqrt{\frac{\pi}{2}} e^{-|x|}\end{aligned}$
    \item $\begin{aligned} \mathfrak{F}^{-1}\left(\frac{-1}{1+4 \alpha^2}\right)(x) & =-\sqrt{\frac{\pi}{2}} \mathfrak{F}^{-1}\left(\sqrt{\frac{2}{\pi}} \frac{1}{1+(2 \alpha)^2}\right)(x) \\ & =-\sqrt{\frac{\pi}{2}} \frac{\frac{1}{2}}{\frac{1}{2}} \mathfrak{F}^{-1}\left(\sqrt{\frac{2}{\pi}} \frac{1}{1+\left(\frac{\alpha}{\frac{1}{2}}\right)^2}\right)(x \mid \\ & =-\sqrt{\frac{\pi}{2}} \frac{1}{2} \mathfrak{F}^{-1}\left(\sqrt{\frac{2}{\pi}} \frac{1}{1+\alpha^2}\right)\left(\frac{x}{2}\right) \\ & =-\frac{1}{2} \sqrt{\frac{\pi}{2}} e^{-\left|\frac{x}{2}\right|}\end{aligned}$
 \end{itemize}
 All together this gives
    $$
    f(x) = \begin{cases}-\frac{1}{2} \sqrt{\frac{\pi}{2}} e^{-\left|\frac{x}{2}\right|}+3 \sqrt{\frac{\pi}{2}} e^{-|x|}+\frac{1}{4} \sqrt{\frac{\pi}{2}} e^{-i \frac{3}{4} x} & -4 \leq x \leq 4 \\ -\frac{1}{2} \sqrt{\frac{\pi}{2}} e^{-\left|\frac{x}{2}\right|}+3 \sqrt{\frac{\pi}{2}} e^{-|x|} & \text { else }\end{cases}
    $$
\end{solution}


















\begin{exercise}
    Draw the complex exponentials $e^{zt}$ in the complex plane for $t=0,1,2,3,4$, 
    where $z$ is one of the following complex numbers:
    \begin{gather*}
        z_1 = \frac 1 2,
        \quad 
        z_2 = - \frac 1 2,
        \quad 
        z_3 = 0.2i,
        \quad 
        z_4 = -0.2i,
        \quad 
        z_3 = - \frac 1 2 + 0.2i
        .
    \end{gather*}
    \textit{You may use a calculater.}
\end{exercise}

\begin{solution}     
\end{solution}

\begin{exercise}
    We consider the Poisson problem with Dirichlet boundary conditions over the interval $[0,L]$:
    \begin{gather*}
        - \Delta u(x) = x^2, \quad 0 < x < L,
        \\
        u(0) = 1, \quad u(L) = 2
    \end{gather*}
    \begin{itemize}
        \item Solve this problem directly. The solution is a polynomial of order $4$.
        \item Extend the right-hand side $f(x) = x^2$ to an odd function with period $2L$ and compute its Fourier coefficients.
        \item How do you use the superposition principle to split the problem? Using these coefficients, find the Fourier series of the solution $u^f$.
    \end{itemize}
\end{exercise}
\begin{solution}     
\end{solution}

\begin{exercise}[Extra]
    Directly compute the solution of the problem
    \begin{gather*}
        - \Delta u(x) = x^2, \quad 0 < x < L,
        \\
        u(0) = 0, \quad u(L) = 0
    \end{gather*}
    and find its Fourier coefficients. Compare this with the function $u^f$ from the previous exercise.
\end{exercise}
\begin{solution}     
\end{solution}

\begin{exercise}
    We solve the Poisson problem with homogeneous Dirichlet boundary conditions over $[0,1]$:
    \begin{gather*}
        - \Delta u(x) = \left\{\begin{array}{ll} x & \text{ if } 0 < x \leq 0.5 \\ 0 & \text{ if } 0.5 < x \leq 1 \end{array}\right., \qquad 0 < x < 1,
        \\
        u(0) = 0, \quad u(1) = 0
    \end{gather*}
    \begin{itemize}
        \item Extend the right-hand side $f(x)$ to an odd function with period $2L$ and compute its Fourier coefficients.
        \item Using these coefficients, find the Fourier series of the solution $u$. Verify that the boundary condition $u(0) = u(L) = 0$ is satisfied.
    \end{itemize}
\end{exercise}
\begin{solution}     
\end{solution}

\begin{exercise}
    We want to find a solution to the boundary value problem 
    \begin{gather*}
        - \Delta u(x) + u(x) = x, \quad 0 < x < L,
        \\
        u(0) = 0, \quad u(L) = 0.
    \end{gather*}
    \begin{itemize}
        \item Extend the right-hand side $f(x) = x$ to an odd function with period $2L$ and compute its Fourier coefficients.
        \item Using these coefficients, find the Fourier series of the solution $u$. Verify that the boundary condition $u(0) = u(L) = 0$ is satisfied.
    \end{itemize}
\end{exercise}
\begin{solution}     
\end{solution}




\begin{exercise}
    Consider the following Fourier sine series for two functions $u, f : \bbR \to \bbR$ with period $T$:
    \begin{align*}
        u(x) = \sum_{n=1}^{\infty} b_{n}^{u} \sin\left( \frac{2\pi n}{T} x \right)
        ,
        \quad 
        f(x) = \sum_{n=1}^{\infty} b_{n}^{f} \sin\left( \frac{2\pi n}{T} x \right)
        .
    \end{align*}
    Express the Fourier coefficients of $f$ by the Fourier coefficients of $u$, and vice-versa, express the Fourier coefficients of $u$ by the Fourier coefficients of $f$,
    in the following cases:
    \begin{enumerate}[label=(\alph*)]
     \item 
     $-u''(x) = f(x)$.
     \item 
     $u''''(x) = f(x)$.
     \item 
     $-u''(x) + 25 u(x) = f(x)$.
     \item 
     $u''''(x) - u''(x) = f(x)$.
     \item 
     $-u''(x) + \gamma \cdot u''(x) = f(x)$.
    \end{enumerate}
    In the last item, $\gamma \in \bbR$ is a constant. For which values of $\gamma$ can you always find a solution?
\end{exercise}
\begin{solution}     
\end{solution}

\begin{exercise}
    Consider the function 
    \begin{align*}
        f : [0,1] \to \bbR, \quad x \mapsto x^{3}.
    \end{align*}
    Extend this to an odd function with period $T = 2$. Sketch the graph of that function from $-2$ to $2$.
    Compute its Fourier coefficients in standard form. Compute the complex Fourier coefficients.
\end{exercise}
\begin{solution}     
\end{solution}




\begin{exercise}
    Given the following functions over an interval $[0,1)$, 
    \begin{enumerate}[label=(\alph*)]
        \item $f(x) = x$
        \item $g(x) = x^2$
        \item $h(x) = e^x$
        \item $s(x) = sin(\pi x)$
    \end{enumerate}
    sketch their extension to 
    \begin{itemize}
        \item a periodic function with period $1$,
        \item an even periodic function with period $2$,
        \item an odd periodic function with period $2$.
    \end{itemize}
    and state the formulas for these functions over $[-1,1]$.
\end{exercise}

\begin{solution}     
\end{solution}



\begin{exercise}[Fun with Neumann boundary conditions 1]
    Consider the Poisson problem with Neumann boundary conditions over the interval $[a,b]$ with constant right-hand side: 
    \begin{gather*}
        - u''(x) = c, \quad a < x < b,
        \\
        u'(a) = h_a, \quad u'(b) = h_b.
    \end{gather*}
    \begin{enumerate}[label=(\alph*)]
        \item
        What is the compatibility condition?
        \item 
        Since $u$ has a constant second derivative, it must be a quadratic function $u(x) = c_1 x^2 + c_2 x + c_3$. 
        Assuming the compatibility condition, find the solution(s).
    \end{enumerate}
\end{exercise}
\begin{solution}     
\end{solution}

\begin{exercise}[Fun with Neumann boundary conditions 2]
    Suppose we have got a Poisson problem with Neumann boundary conditions over the interval $[a,b]$. 
    \begin{gather*}
        - u''(x) = f(x), \quad a < x < b,
        \\
        u'(a) = h_a, \quad u'(b) = h_b.
    \end{gather*}
    We need the compatibility condition 
    \begin{gather*}
        \int_a^b f(x) dx + ( h_b - h_a ) = 0.
    \end{gather*}
    \begin{enumerate}[label=(\alph*)]
        \item
        Find a constant function $c^f : [a,b] \to \bbR$ such that $f - c_f$ has integral zero.
        \item
        Use this to split up the original problem into two subproblems: 
        one with homogeneous boundary conditions, one with constant right-hand side, such that both problems satisfy the compatibility condition.
    \end{enumerate}
\end{exercise}
\begin{solution}     
\end{solution}

\end{document}
