\documentclass[11pt]{article}
\def\hidesolutions{}
\input{header}
\begin{document}
\exsheet{12}{5}{December} % parameters are the number of the session and the day

\begin{exercise}
    Draw the complex exponentials $e^{zt}$ in the complex plane for $t=0,1,2,3,4$, 
    where $z$ is one of the following complex numbers:
    \begin{gather*}
        z_1 = \frac 1 2,
        \quad 
        z_2 = - \frac 1 2,
        \quad 
        z_3 = 0.2i,
        \quad 
        z_4 = -0.2i,
        \quad 
        z_3 = - \frac 1 2 + 0.2i
        .
    \end{gather*}
    \textit{You may use a calculater.}
\end{exercise}

\begin{solution}     
\end{solution}

\begin{exercise}
    Suppose that $f(x) = x^2$ and that
    \[
        g(x) = \left\{\begin{array}{cc} e^{-x} & \text{ if $x > 0$ }, \\ 0 & \text{ otherwise. } \end{array}\right.
    \]
    Compute the convolution $h(x) = (f \star g)(x)$ .
\end{exercise}

\begin{solution}     
\end{solution}

\begin{exercise}
    We consider the Poisson problem with Dirichlet boundary conditions over the interval $[0,L]$:
    \begin{gather*}
        - \Delta u(x) = x^2, \quad 0 < x < L,
        \\
        u(0) = 1, \quad u(L) = 2
    \end{gather*}
    \begin{itemize}
        \item Solve this problem directly. The solution is a polynomial of order $4$.
        \item Extend the right-hand side $f(x) = x^2$ to an odd function with period $2L$ and compute its Fourier coefficients.
        \item How do you use the superposition principle to split the problem? Using these coefficients, find the Fourier series of the solution $u^f$.
    \end{itemize}
\end{exercise}
\begin{solution}     
\end{solution}

\begin{exercise}[Extra]
    Directly compute the solution of the problem
    \begin{gather*}
        - \Delta u(x) = x^2, \quad 0 < x < L,
        \\
        u(0) = 0, \quad u(L) = 0
    \end{gather*}
    and find its Fourier coefficients. Compare this with the function $u^f$ from the previous exercise.
\end{exercise}
\begin{solution}     
\end{solution}

\begin{exercise}
    We solve the Poisson problem with homogeneous Dirichlet boundary conditions over $[0,1]$:
    \begin{gather*}
        - \Delta u(x) = \left\{\begin{array}{ll} 0 & \text{ if } 0 < x < 0.5 \\ 0 & \text{ if } 0.5 < x < 1 \end{array}\right., \qquad 0 < x < 1,
        \\
        u(0) = 0, \quad u(1) = 0
    \end{gather*}
    \begin{itemize}
        \item Extend the right-hand side $f(x) = x$ to an odd function with period $2L$ and compute its Fourier coefficients.
        \item Using these coefficients, find the Fourier series of the solution $u$. Verify that the boundary condition $u(0) = u(L) = 0$ is satisfied.
    \end{itemize}
\end{exercise}
\begin{solution}     
\end{solution}

\begin{exercise}
    We want to find a solution to the boundary value problem 
    \begin{gather*}
        - \Delta u(x) + u(x) = x, \quad 0 < x < L,
        \\
        u(0) = 0, \quad u(L) = 0.
    \end{gather*}
    \begin{itemize}
        \item Extend the right-hand side $f(x) = x$ to an odd function with period $2L$ and compute its Fourier coefficients.
        \item Using these coefficients, find the Fourier series of the solution $u$. Verify that the boundary condition $u(0) = u(L) = 0$ is satisfied.
    \end{itemize}
\end{exercise}
\begin{solution}     
\end{solution}



\end{document}
